Exercising choice over feedback schedules during practice is not advantageous for motor learning

### 2022-05-05 docs/preprint/supplementary.tex
init

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{Self-controlled, Knowledge of results, OPTIMAL theory, Retention}
\usepackage{csquotes}
\usepackage{pdflscape}
\usepackage{setspace}
\raggedbottom
\renewcommand\author[1]{}
\renewcommand\affiliation[1]{}
\authorsnames[1, 1, 1, 1, 1, 1, {2,3}, 1]{Laura St. Germain, Brad McKay, Andrew Poskus, Allison Williams, Olena Leshchyshen, Sherry Feldman, Joshua G.A. Cashaback, Michael J. Carter\vspace{2ex}}
\authorsaffiliations{{Department of Kinesiology, McMaster University}, {Department of Biomedical Engineering, University of Delaware}, {Interdisciplinary Neuroscience Graduate Program, University of Delaware}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Exercising choice over feedback schedules during practice is not advantageous for motor learning},
  pdflang={en-EN},
  pdfkeywords={Self-controlled, Knowledge of results, OPTIMAL theory, Retention},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\authornote{

\vspace{-0.5cm}

\noindent \addORCIDlink{Laura St. Germain}{0000-0002-5513-4183} \newline
\noindent \addORCIDlink{Brad McKay}{0000-0002-7408-2323} \newline
\noindent \addORCIDlink{Joshua G.A. Cashaback}{0000-0002-8642-6648} \newline
\noindent \addORCIDlink{Michael J. Carter}{0000-0002-0675-4271}
\noindent Data and code: \url{https://github.com/cartermaclab/expt_sc-feedback-characteristics} \vspace{2ex} \newline
\noindent \textbf{Corresponding authors:} Michael J. Carter (\href{mailto:michaelcarter@mcmaster.ca}{\nolinkurl{michaelcarter@mcmaster.ca}}) and Laura St.~Germain (\href{mailto:stgerml@mcmaster.ca}{\nolinkurl{stgerml@mcmaster.ca}}).

@@

\abstract{%
The idea that there is a self-controlled learning advantage, where individuals demonstrate improved motor learning after exercising choice over an aspect of practice compared to no-choice groups, has different causal explanations according to the OPTIMAL theory or an information-processing perspective. Within OPTIMAL theory, giving learners choice is considered an autonomy-supportive manipulation that enhances expectations for success and intrinsic motivation. In the information-processing view, choice allows learners to engage in performance-dependent strategies that reduce uncertainty about task outcomes. To disentangle these potential explanations, we provided participants in choice and yoked groups with error or graded feedback (Experiment 1) and binary feedback (Experiment 2) while learning a rapid reaching task with spatial and timing goals. Across both experiments (N = 228 participants), we did not find evidence to support a self-controlled learning advantage. Exercising choice during practice did not increase perceptions of autonomy, competence, or intrinsic motivation, nor did it lead to more accurate error estimation skills. Both error and graded feedback facilitated skill acquisition and learning, whereas no improvements from pre-test performance were found with binary feedback. Finally, the impact of graded and binary feedback on perceived competence highlights a potential dissociation of motivational and informational roles of feedback. Although our results regarding self-controlled practice conditions are difficult to reconcile with either the OPTIMAL theory or the information-processing perspective, they are consistent with a growing body of evidence that strongly suggests self-controlled conditions are not an effective approach to enhance motor performance and learning.
}

The underlying source of errors in skilled actions are often ambiguous and difficult to assign as the learner must rely on noisy and delayed sensory information. Feedback from an external source, such as a coach or computer display, can facilitate or augment this process (Sigrist et al., 2013). Knowledge of results feedback (Salmoni et al., 1984) can provide varying amounts of information to learners depending on its characteristics. Error feedback provides precise information about the magnitude and direction of the error (e.g., -42 cm), graded feedback provides coarse information about either the magnitude or direction of the error (e.g., ``too far''), and binary feedback indicates only success or failure information (e.g., ``miss'') (Luft, 2014). When to provide this feedback is often decided by an external agent; however, this feedback decision can also be made by the learner, a form of self-controlled learning. These self-controlled feedback schedules have typically enhanced motor skill learning compared to yoked feedback schedules, wherein learners experience the feedback schedule created by a self-controlled counterpart, but without any choice (see Sanli et al., 2013; Ste-Marie et al., 2020 for reviews).

Why self-controlled learning advantages emerge has garnered considerable attention in the motor skill learning literature. Within their OPTIMAL (Optimizing performance through intrinsic motivation and attentional learning) theory of motor learning, Wulf and Lewthwaite (2016) have argued that providing participants the opportunity to exercise choice, as in a self-controlled group, creates a virtuous cycle. Specifically, choice leads to increased (perceived) autonomy, leading to enhanced expectancies (e.g., perceived competence) and increased (intrinsic) motivation. These motivational influences lead to improved motor performance, creating a positive feedback loop that ultimately enhances motor learning compared to those not given the same choice opportunities. Support for this view has been drawn from experimental work where participants exercise choice over task-irrelevant or incidental choices. Exercising choice over the color of golf balls to putt (Lewthwaite et al., 2015 Experiment 1) or the mat underneath a target (Wulf et al., 2018 Experiment 1), which picture to hang in a lab (Lewthwaite et al., 2015 Experiment 2), hand order in a maximal force production task (Iwatsuki et al., 2017), which photos to look at while running (Iwatsuki et al., 2018), and the order of exercises to perform (Wulf et al., 2014) have been suggested to improve motor performance or learning. Other research, however, have failed to replicate this benefit of task-irrelevant or incidental choices on motor performance or learning (Carter \& Ste-Marie, 2017a; Grand et al., 2017; McKay \& Ste-Marie, 2020a, 2020b).

Rather than a motivational account, others have forwarded an information-processing explanation. From this perspective, exercising choice allows learners to tailor practice to their individual needs (Chiviacowsky \& Wulf, 2005, 2002) by engaging in performance-contingent strategies (Carter et al., 2014, 2016; Laughlin et al., 2015; Pathania et al., 2019) to reduce uncertainty about movement outcomes (Barros et al., 2019; Carter et al., 2014; Carter \& Ste-Marie, 2017a, 2017b; Grand et al., 2015). Evidence for this view has come from experiments that showed the timing of the feedback decision relative to task performance matters (Carter et al., 2014; Chiviacowsky \& Wulf, 2005), that task-relevant choices are more effective than task-irrelevant choices (Carter \& Ste-Marie, 2017a; cf. Wulf et al., 2018 Experiment 2), that interfering with information-processing activities during (Couvillion et al., 2020; Woodard \& Fairbrother, 2020) or after (Carter \& Ste-Marie, 2017b; Woodard \& Fairbrother, 2020) task performance eliminates self-controlled learning benefits, and that the ability to accurately estimate one's performance is enhanced in choice compared to yoked groups (Carter et al., 2014; Carter \& Patterson, 2012). Thus, further investigation is required to test predictions from these two explanations to better understand why exercising choice during practice confers an advantage for motor skill learning.

To dissociate between the motivational and information-processing accounts of the self-controlled learning advantage, we provided participants in choice and yoked (i.e., no-choice) groups with error or graded feedback (Experiment 1) and binary feedback (Experiment 2) during the acquisition of a rapid aiming task. Motor learning was assessed using delayed (\textasciitilde24 hours) retention and transfer tests. If the OPTIMAL theory is correct, we hypothesized that the characteristics of one's feedback schedule \emph{would not matter} for the self-controlled learning advantage as this advantage arises from the opportunity for choice--a common feature of all choice groups. Thus, we predicted all choice groups would demonstrate superior performance and learning compared to the yoked groups. Alternatively, if the information-processing account is correct, we hypothesized that the characteristics of one's feedback schedule \emph{would matter} for the self-controlled learning advantage as feedback with greater informational value would be more effective for reducing uncertainties about movement outcomes. Thus, we predicted that choice over an error feedback schedule would be the most effective pairing for performance and learning. We also included self-report measures of perceptions of autonomy, competence, and intrinsic motivation, and assessments of error estimation abilities to respectively test auxiliary assumptions of the OPTIMAL theory and information-processing explanations.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study (Simmons et al., 2012). All data and \texttt{R} scripts can be accessed here: \url{https://github.com/cartermaclab/expt_sc-feedback-characteristics}.

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

\hypertarget{experiment-1}{%
\subsubsection{Experiment 1}\label{experiment-1}}

One hundred and fifty-two right-handed (Oldfield, 1971), healthy adults participated in Experiment 1 (\(M_{age}\) = 20.64 years, \(SD_{age}\) = 2.45, 88 females). Sample size was determined from an \emph{a-priori} power analysis using the \emph{ANOVA: fixed effects, main effects and interactions} option in G*Power (Faul et al., 2009) with the following parameters: \(\alpha\) = 0.05, \(\beta\) = .20, \(f\) = 0.23, numerator = 1, and groups = 2. This revealed a required sample of 151 participants. Our chosen effect size was based on a meta-analytic estimate (\(f\) = .32) by McKay et al. (2014); however, we used a more conservative estimate given the uncertainty of how choice would interact with our feedback characteristic manipulation. Participants were compensated \$15 CAD or with course-credit for their time. All participants gave written informed consent and the experiment was approved by McMaster University's Research Ethics Board.

\hypertarget{experiment-2}{%
\subsubsection{Experiment 2}\label{experiment-2}}

A new sample of 76 right-handed (Oldfield, 1971), healthy adults participated in Experiment 2 (\(M_{age}\) = 20.18 years, \(SD_{age}\) = 3.18, 47 females). Sample size was selected so group size matched that used in Experiment 1. Participants were compensated \$15 CAD or with course-credit for their time. All participants gave written informed consent and the experiment was approved by McMaster University's Research Ethics Board.

\hypertarget{task}{%
\subsection{Task}\label{task}}

In Experiments 1 and 2, participants sat in a chair facing a monitor (1920x1080 resolution) with their left arm in a custom manipulandum that restricted movement to the horizontal plane. Their elbow was bent at approximately 90° and they grasped a vertical handle with their left hand. Handle position was adjusted as needed to ensure the central axis of rotation was about the elbow. The task required a rapid ``out-and-back'' movement such that the reversal happened at \(40^\circ\) (in pre-test, acquisition, and retention) or \(60^\circ\) (in transfer). The starting point for all trials was \(0^\circ\). Participants were instructed to make a smooth movement to the reversal and back without hesitating when reversing their movement. The movement time goal to the reversal was always 225 ms. The task and instructions were similar to those used by Sherwood (1996; 2009). Vision of the manipulandum and limb were occluded during all phases of the experiment. Angular displacement for the elbow was collected via a potentiometer attached to the axis of rotation of the custom manipulandum. Potentiometer data were digitally sampled at 1000 Hz (National Instruments PCIe-6321) using a custom LabVIEW program and stored for offline analysis.

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

\hypertarget{experiment-1-1}{%
\subsubsection{Experiment 1}\label{experiment-1-1}}

The first 76 participants were randomly assigned to either the Choice+Error-Feedback group (\(n\) = 38; \(M_{age}\) = 20.24 years, \(SD_{age}\) = 2.37, 22 females) or the Choice+Graded-Feedback group (\(n\) = 38; \(M_{age}\) = 20.76 years, \(SD_{age}\) = 3.02, 26 females). This is typical in the self-controlled learning literature as the self-controlled participants' self-selected feedback schedules are required for providing feedback to the participants in the yoked (i.e., control) groups. The remaining 76 participants were randomly assigned to either the Yoked+Error-Feedback group (\(n\) = 38; \(M_{age}\) = 20.53 years, \(SD_{age}\) = 2.13, 23 females) or the Yoked+Graded-Feedback group (\(n\) = 38; \(M_{age}\) = 21.03 years, \(SD_{age}\) = 2.32, 22 females).

Data collection consisted of two sessions separated by approximately 24 hours.\footnote{Six participants (three Choice+Error-Feedback and three Choice+Graded-Feedback) had their second session completed approximately 48 hours later because a snowstorm closed the University.} Session one included a pre-test (12 trials) and an acquisition phase (72 trials). Session two included the delayed retention (12 trials) and transfer (12 trials) tests. No feedback about motor performance was provided in pre-test, retention, or transfer. Prior to the pre-test, all participants received instructions about the task and its associated spatial and timing goals. Additionally, half of the participants in each group were randomly selected to estimate their performance on the spatial and timing goals after each trial in the pre-test. Only a subset of participants were asked to estimate their performance in pre-test to mitigate the potential that doing so would prompt participants to adopt this strategy during the experiment as error estimation has been suggested (e.g., Chiviacowsky \& Wulf, 2005) to be adopted spontaneously by participants controlling their feedback schedule. However, asking participants to estimate their performance during pre-test is necessary to be able to assess how this skill develops as a function of one's practice condition.

Participants were reminded of the instructions about the task and its associated goals at the start of the acquisition phase. Group specific instructions regarding feedback were also provided. Participants in the Choice+Error-Feedback group and the Choice+Graded-Feedback group were told they could choose their feedback schedule, with the restriction that they must select feedback on 24 of the 72 acquisition trials. They were informed that if the number of remaining feedback requests equaled the number of remaining acquisition trials, these trials would default to feedback trials. This feedback restriction was implemented to ensure the relative frequency of feedback was equated across all groups. Similar restrictions have been used in past research involving multiple choice groups (e.g., Chiviacowsky \& Wulf, 2005). Participants in the Yoked+Error-Feedback group and the Yoked+Graded-Feedback group were told they may or may not receive feedback following a trial based on a predetermined schedule. Thus, participants in these groups were not aware that their feedback schedule was actually created by a participant in a corresponding choice group. While this yoking procedure ensures that the total number of feedback trials and their relative placement during acquisition are identical, the content of the feedback reflected each participant's own performance. Error feedback for the spatial and timing goals was provided as the difference between the participant's actual performance and the task goal (i.e., constant error). Graded feedback for the spatial goal was provided as ``too short'' if performance was \textless{} 40 degrees (or 60 degrees in transfer), ``hit'' if exactly 40 degrees, and ``too far'' if \textgreater{} 40 degrees. For the timing goal, graded feedback was provided as ``too fast'' when performance was \textless{} 225 ms, ``hit'' if exactly 225 ms, and ``too slow'' if \textgreater{} 225 ms. All participants were shown a sample feedback display that corresponded to their experimental group and were asked to interpret it aloud for the researcher to verify understanding.

@@

{\centering \includegraphics[height=0.45\textheight]{../../figs/fig1}

@@

\caption{\small \onehalfspacing \textbf{Overview of a typical acquisition trial for the choice groups.} The sequence of events a participant in the choice groups experienced during the acquisition phase. Trials began by informing participants the trial number (500 ms) they were on in acquisition. Shortly after, the text ``Get Ready!'' appeared on the screen and 800 ms later a visual go-signal was presented in the form of a green circle in the center of the screen. Participants began their movement when ready after seeing the visual go-signal as we were not interested in reaction time. While participants completed their rapid out-and-back movement, the computer screen was blank. Upon returning to the starting position, a red circle appeared in the center of the screen. A 2000 ms feedback delay interval was used and this interval was followed by the feedback prompt. The feedback prompt also displayed an updated counter representing the number of feedback trials they had left. If the number of remaining feedback trials matched the number of acquisition trials left, these trials automatically defaulted to feedback trials. On trials where feedback was not requested, a blank screen \textbf{(A)} was shown for 3000 ms. When feedback was selected, feedback was provided for both the spatial and timing goals according to their experimental group. The error feedback group \textbf{(B)} saw their constant error, the graded feedback group \textbf{(C)} saw either ``too far'' or ``too short'' for the spatial goal and ``too fast'' or ``too slow'' for the timing goal, and the binary feedback group \textbf{(D)} saw either ``hit'' or ``miss'' for the task goals. The sequence of events was the same for the yoked groups with the exception they did not see a feedback prompt. The sequence of events was similar in pre-test, retention, and transfer except all trials were no-feedback trials.}\label{fig:fig1}
\end{figure}

A typical acquisition trial (see Figure \ref{fig:fig1}) began with the current trial number displayed (500 ms), followed by a visual ``Get Ready!'' and a visual go-signal (800 ms apart). Participants were free to begin their movement when ready following the visual go-signal (i.e., green circle) as this was not a reaction time task. The computer screen was blank while participants made their movement. When participants returned to the starting position, a red circle was displayed on the monitor. Following a 2000 ms feedback delay interval, the feedback decision prompt was presented for the self-controlled groups. If feedback was not selected, a blank screen was displayed for 3000 ms. If feedback was selected (or imposed on the yoked groups), it was also displayed for 3000 ms.

Before the retention and transfer tests, participants were reminded about the task and its associated goals. All participants were asked to estimate their performance after each trial in retention and transfer. After the pre-test, trials 12 and 72 in acquisition, and before the delayed retention test, participants answered a series of questions pertaining to perceived competence, task interest and enjoyment, and perceived autonomy. The perceived competence and task interest and enjoyment questions were from the Intrinsic Motivation Inventory (McAuley et al., 1989; Ryan, 1982) and the perceived autonomy questions were used in earlier work (Barros et al., 2019; Carter \& Ste-Marie, 2017a). Cronbach's alpha values for each questionnaire at each time point are reported in Table \ref{tab:table1}.

\hypertarget{experiment-2-1}{%
\subsubsection{Experiment 2}\label{experiment-2-1}}

Similar to Experiment 1, the first half of participants were assigned to the Choice+Binary-Feedback group (\(n\) = 38; \(M_{age}\) = 22.37 years, \(SD_{age}\) = 3.13, 19 females) and the remaining participants were assigned to the Yoked+Binary-Feedback group (\(n\) = 38; \(M_{age}\) = 18.00 years, \(SD_{age}\) = 0.93, 28 females). Binary feedback for the spatial goal was provided as ``hit'' if performance was exactly 40 degrees (or 60 degrees in transfer) and as ``miss'' for everything else. For the timing goal, binary feedback was provided as ``hit'' when performance was exactly 225 ms and as ``miss'' for everything else. Data collection was identical to that of Experiment 1, except in the acquisition instructions participants in both groups were shown a sample binary feedback display and were asked to interpret it aloud for the researcher to verify understanding.

\begin{table}

\caption{\label{tab:table1}Cronbach's alpha for each questionnaire at each timepoint.}
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{12em}>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{6em}>{\raggedright\arraybackslash}p{7em}}
\toprule
Questionnaire & After pre-test & After block 1 & After block 6 & Before retention\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Experiment 1}}\\
\hspace{1em}Perceived autonomy & 0.68 & 0.81 & 0.80 & 0.83\\
\hspace{1em}Perceived competence & 0.86 & 0.93 & 0.95 & 0.94\\
\hspace{1em}Intrinsic motivation & 0.86 & 0.90 & 0.92 & 0.93\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Experiment 2}}\\
\hspace{1em}Perceived autonomy & 0.39 & 0.73 & 0.79 & 0.85\\
\hspace{1em}Perceived competence & 0.92 & 0.88 & 0.91 & 0.91\\
\hspace{1em}Intrinsic motivation & 0.91 & 0.91 & 0.93 & 0.94\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}\textit{Note.} Block 1 and 6 are from the acquisition phase.}\\
\end{tabular}
\end{table}

\hypertarget{data-analysis}{%
\subsection{Data Analysis}\label{data-analysis}}

Movement trajectories for all trials were visually inspected by a researcher and trials with errors (e.g., technical issues, moving before the ``go'' signal) were removed. A total of 4.03\% (662/16146) and 3.73\% (306/8208) of trials for Experiments 1 and 2 were removed, respectively. Trials were aggregated into blocks of 12 trials, resulting in one block of trials for pre-test, retention, and transfer, and six blocks of trials for acquisition. Our primary performance outcome variable was total error (E) (Henry, 1974, 1975) and was computed using the equation:
\begin{equation}
E = \sqrt{\sum{(x_{i} - T)^2 / n}}
\end{equation}
where \(x_{i}\) is the score on the \(i\)th trial, \(T\) is the target goal, and \(n\) is the number of trials in a block.

To test for performance differences in pre-test, retention, and transfer, total error for the spatial and timing goals were analyzed in separate mixed ANOVAs (\emph{Experiment 1:} 2 Choice x 2 Feedback x 3 Test; \emph{Experiment 2:} 2 Choice x 3 Test). To test for performance differences during acquisition, total error for the spatial and timing goals during acquisition were analyzed in separate mixed ANOVAs (\emph{Experiment 1:} 2 Choice x 2 Feedback x 6 Block; \emph{Experiment 2:} 2 Choice x 6 Block). Our primary psychological outcome variables were intrinsic motivation (i.e., interest/enjoyment), perceived competence, and perceived autonomy. The mean score of the responses for these constructs at each time point was calculated for each participant and analyzed in separate mixed ANOVAs (\emph{Experiment 1:} 2 Choice x 2 Feedback x 4 Time; \emph{Experiment 2:} 2 Choice x 4 Time). Of secondary interest, error estimation abilities were assessed as total error between a participant's estimation and actual performance in pre-test (50\% of the participants in each group in Experiments 1 and 2), retention, and transfer (see \textbf{Supplementary A}).

Alpha was set to .05 for all statistical analyses. Corrected degrees of freedom using the Greenhouse-Geisser technique are always reported for repeated measures with more than two levels. Generalized eta squared \((\eta_{G}^2)\) is provided as an effect size statistic (Bakeman, 2005; Lakens, 2013; Olejnik \& Algina, 2003) for all omnibus tests. Post hoc comparisons were Holm-Bonferroni corrected to control for multiple comparisons. Statistical tests were conducted using R (Version 4.1.2; R Core Team, 2021) and the R-packages \emph{afex} (Version 1.1.1; Singmann et al., 2021), \emph{computees} (Re, 2013), \emph{ggResidpanel} (Version 0.3.0.9000; Goode \& Rey, 2022), \emph{kableExtra} (Version 1.3.4; Zhu, 2021), \emph{metafor} (Version 3.4.0; Viechtbauer, 2010), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{patchwork} (Version 1.1.0.9000; Pedersen, 2020), \emph{rogme} (Version 0.2.1; Rousselet et al., 2017), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), and \emph{tinylabels} (Version 0.2.3; Barth, 2022) were used in this project.

\hypertarget{results}{%
\section{Results}\label{results}}

Model diagnostics of total error for the spatial and timing goals revealed skewed distributions. We therefore conducted sensitivity analyses using the shift function, which is a robust statistical method well-suited for skewed distributions (Rousselet \& Wilcox, 2020; Wilcox, 2021). The results of these analyses (see \textbf{Supplementary B}) were consistent with those of the mixed ANOVAs, which we report below.

\hypertarget{pre-test-retention-and-transfer}{%
\subsection{Pre-test, retention, and transfer}\label{pre-test-retention-and-transfer}}

\hypertarget{experiment-1-2}{%
\subsubsection{Experiment 1}\label{experiment-1-2}}

Spatial (Fig. \ref{fig:fig2}A) and timing (Fig. \ref{fig:fig2}B) error decreased from the pre-test to the retention and transfer tests. There was a main effect of Test for spatial error, \(F(1.33,196.52) = 40.20\), \(p < .001\), \(\eta_{G}^2 = .138\), where performance was less errorful in retention and transfer than pre-test (\(p\)'s \textless{} .001) and performance in retention was better than transfer (\(p\) \textless{} .001). A main effect of Test was also found for timing error, \(F(1.08,160.23) = 81.21\), \(p < .001\), \(\eta_{G}^2 = .245\), with pre-test performance more errorful than both retention and transfer (\(p\)'s \textless{} .001), and retention was less errorful than transfer (\(p\) \textless{} .001).

\hypertarget{experiment-2-2}{%
\subsubsection{Experiment 2}\label{experiment-2-2}}

Spatial (Fig. \ref{fig:fig3}A) and timing (Fig. \ref{fig:fig3}B) error did not change considerably from the pre-test to the retention and transfer tests. There were no significant main effects or interactions for either task goals.

\vspace{4ex}

@@

{\centering \includegraphics{../../figs/fig2}

@@

\caption{\small \onehalfspacing \textbf{Experiment 1 data.} The Choice with error feedback (Choice+Error) group is shown in dark blue circles, the Choice with graded feedback (Choice+Graded) group is shown in light blue squares, the Yoked with error feedback (Yoked+Error) group is shown in red triangles, and the Yoked with graded feedback (Yoked+Graded) group is shown in yellow crosses. Error bars denote 95\% bootstrapped confidence intervals. \textbf{(A)} Spatial total error (degrees) and \textbf{(B)} timing total error (ms) averaged across blocks and participants within each group. Dotted vertical lines denote the different experimental phases. Pre-test and acquisition occurred on Day 1 and retention and transfer occurred approximately 24-hours later on Day 2. Self-reported scores for perceived autonomy \textbf{(C)}, perceived competence \textbf{(D)}, and intrinsic motivation \textbf{(E)} after the pre-test and after blocks 1 and 6 of acquisition on Day 1, and before the retention test on Day 2. Scores could range on a Likert scale from 1 (Strongly disagree) to 7 (Strongly agree). Dots represent individual data points.}\label{fig:fig2}
\end{figure}

@@

{\centering \includegraphics{../../figs/fig3}

@@

\hypertarget{acquisition}{%
\subsection{Acquisition}\label{acquisition}}

\hypertarget{experiment-1-3}{%
\subsubsection{Experiment 1}\label{experiment-1-3}}

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\usepackage{csquotes}
\usepackage{pdflscape}
\usepackage{setspace}
\raggedbottom
\AtBeginDocument{\let\maketitle\relax}
\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Exercising choice over feedback schedules during practice is not advantageous for motor learning},
  pdflang={en-EN},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\hypertarget{supplementary-a}{%
\section{Supplementary A}\label{supplementary-a}}

We asked all participants to estimate their performance on the spatial (Fig. \ref{fig:figS1}) and timing (Fig. \ref{fig:figS2}) components of the motor task after each trial in retention and transfer, similar to past research (e.g., Barros et al., 2019; Carter et al., 2014; Carter \& Patterson, 2012). Additionally, half of the participants in each group were randomly selected to also estimate their performance after each trial in the pre-test (not shown). To assess error estimation for the spatial and timing components, we first calculated the difference between a participant's estimated performance and their actual performance on each trial. Next, we computed total estimation error (EE) using the equation:
\begin{equation}
EE = \sqrt{CE^2 + VE^2}
\end{equation}
where \(CE\) was the average estimation bias and \(VE\) was the standard deviation of these errors. This approach is consistent with that used by Bruechert et al. (2003).

\hypertarget{error-estimation-in-retention-and-transfer}{%
\subsection{Error estimation in retention and transfer}\label{error-estimation-in-retention-and-transfer}}

Total estimation error was generally lower in retention compared to transfer for both the spatial (Fig. \ref{fig:figS1}) and timing (Fig. \ref{fig:figS2}) domains in Experiments 1 and 2. For the spatial estimation error, we found a significant main effect of Test in Experiment 1, \(F(1,148) = 32.40\), \(p < .001\), \(\eta_{G}^2 = .060\), and in Experiment 2, \(F(1,74) = 18.79\), \(p < .001\), \(\eta_{G}^2 = .026\), with more accurate estimations in retention than transfer. Although there was a significant Choice x Feedback x Test interaction in Experiment 1 for timing estimation error, \(F(1,148) = 7.12\), \(p = .008\), \(\eta_{G}^2 < .001\), we did not decompose the interaction as the effect size estimate was less than \(.001\). There were no significant main effects or interactions for timing estimation error in Experiment 2. A potential explanation for the difference in estimation accuracy between retention and transfer for the spatial component and not the the timing component is that only the spatial goal changed. Specifically, it was 40 deg in retention and 60 deg in transfer whereas the timing goal remained 225 ms in both tests.

{\centering \includegraphics{../../figs/figS1}

@@

\caption{\small \onehalfspacing \textbf{Total spatial error estimation data.} Boxplots of retention and transfer data from all participants for Experiment 1 \textbf{(A)} and Experiment 2 \textbf{(B)}. The Choice with error feedback (Choice+Error) group is shown in dark blue, the Choice with graded feedback (Choice+Graded) group is shown in light blue, the Choice with binary feedback (Choice+Binary) group is shown in green, the Yoked with error feedback (Yoked+Error) group is shown in red, the Yoked with graded feedback (Yoked+Graded) group is shown in yellow, and the Yoked with binary feedback (Yoked+Binary) group is shown in purple. Boxplots represent 25th, 50th, and 75th percentiles, and the solid black line denotes the group mean. Grey connected dots represent individual data for participants in each group. Individual data of 1 participant in each of the Choice+Error, Yoked+Error, and Yoked+Binary groups is not shown as their error exceeded 500 ms.}\label{fig:figS1}
\end{figure}

@@

{\centering \includegraphics{../../figs/figS2}

@@

\caption{\small \onehalfspacing \textbf{Total timing error estimation data.} Boxplots of retention and transfer data from all participants for Experiment 1 \textbf{(A)} and Experiment 2 \textbf{(B)}. The Choice with error feedback (Choice+Error) group is shown in dark blue, the Choice with graded feedback (Choice+Graded) group is shown in light blue, the Choice with binary feedback (Choice+Binary) group is shown in green, the Yoked with error feedback (Yoked+Error) group is shown in red, the Yoked with graded feedback (Yoked+Graded) group is shown in yellow, and the Yoked with binary feedback (Yoked+Binary) group is shown in purple. Boxplots represent 25th, 50th, and 75th percentiles, and the solid black line denotes the group mean. Grey connected dots represent individual data for participants in each group.}\label{fig:figS2}
\end{figure}

\hypertarget{development-of-error-estimation-skills}{%
\subsection{Development of error estimation skills}\label{development-of-error-estimation-skills}}

In Experiment 1, the subset of participants in each group who estimated their error after each pre-test trial improved the accuracy of their estimation skills throughout the experiment. We found significant main effects of Test for both the spatial, \(F(1.45,104.67) = 20.72\), \(p < .001\), \(\eta_{G}^2 = .144\), and timing, \(F(1.10,79.40) = 46.88\), \(p < .001\), \(\eta_{G}^2 = .060\), error estimations. For the spatial error estimations, Holm-Bonferonni post-hoc tests revealed that pre-test had higher error than retention and transfer (\(p\)'s \textless{} .001), and transfer had higher error than retention (\(p\) \textless{} .001). Similarly, timing error estimations were more accurate in retention and transfer compared to the pre-test (\(p\)'s \textless{} .001). In Experiment 2, none of the main effects or interactions were significant.

\hypertarget{supplementary-b}{%
\section{Supplementary B}\label{supplementary-b}}

Model diagnostics of total error (E) for the spatial and timing goals revealed skewed distributions. As a result, we carried out sensitivity analyses using shift functions, which is a robust technique that is well-suited for skewed distributions (Rousselet \& Wilcox, 2020; Wilcox, 2021). For all shift functions, we collapsed across retention and transfer data to compute a single post-test score for both the spatial and timing goals. Given our primary interest was related to the role of choice during practice for motor learning and shift functions only compare two groups, we only ran these analyses on the choice factor for Experiment 1 (collapsed across feedback) and Experiment 2. Conducting a shift function analysis is a multi-step process that first involved calculating the 20\% trimmed means for each participant and time point. Next, deciles for the factor of choice (choice versus yoked) in Experiments 1 and 2 were calculated using the Harrell-Davis estimator (Harrell \& Davis, 1982; Rousselet et al., 2017). Lastly, differences at each decile were evaluated based on the 95\% confidence intervals corrected for multiple comparisons using Hochberg's method (Hochberg, 1988). Spatial and timing measures were analyzed separately for Experiments 1 and 2.

\hypertarget{choice-versus-yoked}{%
\subsection{Choice versus yoked}\label{choice-versus-yoked}}

For Experiment 1, the 95\% confidence intervals overlapped with zero at each decile when comparing the choice and yoked groups for both the spatial (Fig. \ref{fig:figS3}A) and timing (Fig. \ref{fig:figS3}B) goals. Similarly, in Experiment 2 there were no significant differences between the choice and yoked groups in any decile for the spatial (Fig. \ref{fig:figS3}C) and timing (Fig. \ref{fig:figS3}D) goals. The results from these shift function analyses are consistent with the analyses reported in the main manuscript and a recent meta-analysis (McKay et al., in-press), suggesting that giving learners choice during practice is not advantageous for motor learning.

@@

{\centering \includegraphics{../../figs/figS3}

@@

\caption{\small \onehalfspacing \textbf{Post-test (collapsed across retention and transfer) spatial E (deg) shift function.} The top row illustrates scatter plots of individual mean spatial E for each group in Experiment 1 \textbf{(A)} and Experiment 2 \textbf{(B)}. The middle row illustrates the same scatter plots as the top row, but with the deciles of each distribution represented by the black lines. The thick black line represents the the median of each distribution. The deciles from each group are joined by colored lines, with blue (Experiment 1) and green (Experiment 2) indicating lower error for the choice group deciles, and red (Experiment 1) and green (Experiment 2) indicating lower error for the yoked group deciles. The bottom row illustrates the shift function, which focuses on the grey shaded region of the x-axis in the middle row. The deciles for the choice group are plotted on the x-axis and the difference in deciles between the choice and yoked group are plotted on the y-axis. The vertical dash line represents the median of the choice distribution. The same color coding for differences in deciles from the middle row is used in the bottom row. Error bars represent 95\% confidence intervals, corrected for multiple comparisons. The horizontal dashed line represents zero differences between the deciles of the two group. If a 95\% confidence interval overlaps with zero, there is no significant difference between the two groups on that decile of the distribution. There were no significant differences for any decile in either experiment.}\label{fig:figS3}
\end{figure}

@@

{\centering \includegraphics{../../figs/figS4}

@@
